### 2022-02-21 data/processing_data/CountingLines.Rmd

---
title: "RMarkdown lengths"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

This document shows how I generated the dataset `lablines.csv`.

## Step 1: Get the names of all the RMarkdown files

This file uses the `here` package to create paths, but refers to a directory on my local computer that does not appear in the repository. Therefore, this code is not fully reproducible.

```{r}
tidyfolder <- here("..", "..", "STAT220", "LL\ labs", "tidy", "pre-labs")
formulafolder <- here("..", "..", "STAT220", "LL\ labs", "formula", "pre-labs")
tidyfiles <- list.files(path = tidyfolder, full.names=TRUE)[-1] 
names(tidyfiles) <- list.files(path = tidyfolder)[-1]
formulafiles <- list.files(path = formulafolder, full.names=TRUE )[-1]
names(formulafiles) <- list.files(path = formulafolder)[-1]
```

## Step 2: Read the RMarkdown documents

```{r}
tidyfiles_full <- tidyfiles %>%
  map(read_lines) %>%
  map_int(length) %>%
  as_tibble(rownames = "file") %>%
  rename(lines = value)
formulafiles_full <- formulafiles %>%
  map(read_lines) %>%
  map_int(length) %>%
  as_tibble(rownames = "file") %>%
  rename(lines = value)
```

## Step 3: Join the formula and tidyverse data

```{r}
tidyfiles_full <- tidyfiles_full %>%
  mutate(section = "tidy") %>%
  mutate(file = str_remove(file, "STAT_220_"), 
         file = str_remove(file, ".Rmd"), 
         file = str_remove(file, "PreLab"),
         file = str_remove(file, "tidy"),
         file = str_remove(file, "-"),
         file = str_remove(file, "_"))
formulafiles_full <- formulafiles_full %>%
  mutate(section = "formula") %>%
  mutate(file = str_remove(file, "STAT_220_"), 
         file = str_remove(file, ".Rmd"), 
         file = str_remove(file, "PreLab"),
         file = str_remove(file, "formula"),
         file = str_remove(file, "-"),
         file = str_remove(file, "_"))
labs <- tidyfiles_full %>%
  bind_rows(formulafiles_full) 
```

## Step 4: Write out data
```{r}
write_csv(labs, here("data", "processed", "lablines.csv"))
```